First paper - Draft

### 2017-06-04 paper/paper.Rnw
Played around with the Rnw file but wasn't getting it -- moved to only code file first

    df_lower <- filter(df, new_hisced %in% c(1, 2))
    df_upper <- filter(df, new_hisced == 7)

@@

<<modeling>>=
# adapted_year_data <-
#     map(pisa_all2$value, ~ {
#       if (unique(.x$wave) == "pisa2000") {
#         # pisa2000 has a different coding so here I recode 6 to 7 so that in all waves the top edu
#         # is 7 and the bottom is 1
#         .x <- 
#           mutate(.x, new_hisced = as.character(dplyr::recode(as.numeric(high_edu_broad), `6` = 7)))
#       } else {
#         .x <-
#           mutate(.x, new_hisced = as.character(high_edu_broad))
#       }
#       .x
# })
# 
# results_math <- test_diff(adapted_year_data, reliability_pisa, "MATH")
# results_read <- test_diff(adapted_year_data, reliability_pisa, "READ")
# US is missing for reading

# Cache is not working properly for the code above, so I just load the saved cached file
load("./paper/cache/modeling_9a0b38d1d53fa243b0242580f0672fa5.RData")

@@

<<>>=
i <- 3
vals <- intersect(unique(adapted_year_data[[1]]$country), unique(adapted_year_data[[2]]$country))

while (i < 7) {
  vals <- intersect(vals, unique(adapted_year_data[[i]]$country))
  i <- i + 1
}

subset_cnts <- map(adapted_year_data, ~ filter(.x, country %in% vals))

cnt_to_bind <-
  map(subset_cnts, function(df) {

@@

m <- left_join(sample_size, high_low_sample) %>%
  left_join(ses_sample) %>%
  arrange(country, year)

@@

<<graphin differences, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
  reduced_data %>%
    filter(country %in% c("United States")) %>%
    ggplot(aes(as.character(wave), Mean, group = escs_dummy, colour = escs_dummy)) +
    geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
    geom_line() +
    facet_wrap(~ country)
@

@@

reduced_data %>%
    select(country, escs_dummy, Mean, wave) %>%
    spread(key = escs_dummy, value = Mean) %>%
    mutate(diff = `1` - `0`) %>%
    gather(escs_dummy, Mean, -country, -wave, -`1`, -`0`) %>%
    left_join(se_data) %>%
    mutate(continent = countrycode(country, "country.name", "continent"),
           region = countrycode(country, "country.name", "region"),
           lower = Mean - 1.96 * se_diff,
           upper = Mean + 1.96 * se_diff) %>%
    filter(!is.na(continent), !is.na(region), country %in% countries) %>%
    ggplot(aes(as.factor(wave), Mean, group = country, colour = country)) +
    geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1) +
    geom_line() +
    geom_point(size = 0.5) +
    scale_colour_discrete(guide = F) +
    # geom_smooth(aes(group = region), colour = "blue", method = "lm") +
    coord_cartesian(ylim = c(0, 4)) +
    facet_wrap(~ country) +
    ggtitle("math")

>>>>

    df_lower <- filter(df, new_hisced %in% 1)
    df_upper <- filter(df, new_hisced == 7)

@@

<<modeling>>=
adapted_year_data <-
    map(pisa_all2$value, ~ {
      if (unique(.x$wave) == "pisa2000") {
        # pisa2000 has a different coding so here I recode 6 to 7 so that in all waves the top edu
        # is 7 and the bottom is 1
        .x <-
          mutate(.x, new_hisced = as.character(dplyr::recode(as.numeric(high_edu_broad), `6` = 7)))
      } else {
        .x <-
          mutate(.x, new_hisced = as.character(high_edu_broad))
      }
      .x
})

results_math <- test_diff(adapted_year_data, reliability_pisa, "MATH")
# results_read <- test_diff(adapted_year_data, reliability_pisa, "READ")
# US is missing for reading

# Cache is not working properly for the code above, so I just load the saved cached file
# load("./paper/cache/modeling_9a0b38d1d53fa243b0242580f0672fa5.RData")

@@

<<>>=
vals <- c("Australia",
          "Germany",
          "Denmark",
          "Spain",
          "France",
          "Italy",
          "Netherlands",
          "Sweden",
          "United States")
# i <- 3
# vals <- intersect(unique(adapted_year_data[[1]]$country), unique(adapted_year_data[[2]]$country))

# while (i < 7) {
#   vals <- intersect(vals, unique(adapted_year_data[[i]]$country))
#   i <- i + 1
# }

sample_size_calc <- function(df, selected = F, cnts) {

if (selected) df <- map(df, ~ filter(.x, country %in% cnts))

cnt_to_bind <-
  map(df, function(df) {

@@

selected_cnts <-
  left_join(sample_size, high_low_sample) %>%
  left_join(ses_sample) %>%
  arrange(year, country)

selected_cnts
}

selected_cnts <- sample_size_calc(adapted_year_data, vals)  
xtable::xtable(selected_cnts)
@

<<>>=
all_cnts <- sample_size_calc(adapted_year_data)
@

<<>>=
selected_cnt <- map(adapted_year_data, ~ select(.x, country, high_edu_broad, new_hisced, escs_dummy))

selected_cnts <-
  left_join(sample_size, high_low_sample) %>%
  left_join(ses_sample) %>%
  arrange(year, country)

@@

<<graphin differences, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
reduced_data %>%
    filter(country %in% c("United Kingdom")) %>%
    ggplot(aes(as.character(wave), Mean, group = escs_dummy, colour = escs_dummy)) +
    geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
    geom_line() +
    facet_wrap(~ country)
@

@@

reduced_data %>%
    select(country, escs_dummy, Mean, wave) %>%
    spread(key = escs_dummy, value = Mean) %>%
    mutate(diff = `1` - `0`) %>%
    gather(escs_dummy, Mean, -country, -wave, -`1`, -`0`) %>%
    left_join(se_data) %>%
    mutate(continent = countrycode(country, "country.name", "continent"),
           region = countrycode(country, "country.name", "region"),
           lower = Mean - 1.96 * se_diff,
           upper = Mean + 1.96 * se_diff) %>%
    filter(!is.na(continent), !is.na(region), country %in% vals) %>%
    ggplot(aes(as.factor(wave), Mean, group = country, colour = country)) +
    geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1) +
    geom_line() +
    geom_point(size = 0.5) +
    scale_colour_discrete(guide = F) +
    # geom_smooth(aes(group = region), colour = "blue", method = "lm") +
    coord_cartesian(ylim = c(0, 4)) +
    facet_wrap(~ country) +
    ggtitle("math")
